---
title: 'Data Analytics Project 4: Bank Marketing'
author: "Ananya Kaushik, Natasha Pirani, Nolan Bentley, and Juan Ramos Fuentes"
resource_files:
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
require(MASS)
require(ISLR)
require(ggplot2)
require(data.world)
require(plotly)
require(dplyr)
require(leaps)
require(tidyverse)
require(shiny)
require(class)
require(e1071)
require(tree)
require(randomForest)
knitr::opts_chunk$set(echo = TRUE)
options(shiny.sanitize.errors = FALSE)
```

# **Setting Up the Environment**  
## **R Session Info**
```{r, echo=FALSE, message=FALSE, warning=FALSE}
sessionInfo()
```


## **Connecting to data.world**
https://data.world/natashapirani96/f-17-eda-project-4
```{r, echo=FALSE, message=FALSE, warning=FALSE}
source("data.R", echo = FALSE)
```

# **Introduction** 
This document contains an analysis on the marketing campaigns of a Portuguese banking institution that used phone calls.

## **Dataset -  Bank Marketing**
```{r, warning=FALSE}
#source("analysis.R", echo = FALSE)
renderDataTable( options = list(pageLength = 10), {
  bank
})
```

# **Statistical Analysis**
## **Decision Trees**

We want to identify the variables that accurately predict the output variable: did the client subscribe to a term deposit during the direct marketing campaign? We use a tree-based approach: decision trees.
```{r, warning = FALSE, message = FALSE}
bank$output = as.factor(bank$y)
na.omit(bank)
bank = bank %>% mutate_if(is.character, as.factor)
attach(bank)
tree.bank = tree(output ~ . - y, data = bank)
summary(tree.bank)
renderPlot({
  plot(tree.bank)
  text(tree.bank, pretty = 0)
})
tree.bank
```
We first use the entire dataset to fit the tree model. The variables actually used in tree construction were: nr_employed, duration, pdays, cons_conf_idx, and euribor3m.

We then used a random sample from the original dataset and we train the model on that training sample
```{r, message=FALSE, warning=FALSE}
set.seed(1011)
train = sample(1:nrow(bank), 2000)
tree.bank = tree(output ~ . - y, bank, subset = train)
renderPlot({
  plot(tree.bank);text(tree.bank, pretty = 0)
})
```
When we look at the confusion matrix, we get a ~90% accuracy rate
```{r, message=FALSE, warning=FALSE}
tree.pred = predict(tree.bank, bank[-train,], type="class")
with(bank[-train,], table(tree.pred, output))
(1855+72)/2119
```

##**Support Vector Machines**

Here, a model was created using the output (target) variable "y", duration (of the phone call), and consumer confidence index (CCI). This variable "y" is a boolean, indicating whether or not the client has subscribed to the term deposit.

```{r, message=FALSE, warning=FALSE}
bank = na.omit(bank)
df = data.frame(output = as.factor(bank$y), CCI = bank$cons_conf_idx, duration = bank$duration)
svmfit=svm(output~.,df,kernel="linear",cost=1000,scale=FALSE)
svmfit
```

The plot of the linear support vector machine was created:

```{r, message=FALSE, warning=FALSE}
renderPlot({
  plot(svmfit, df, CCI ~ duration)
})
```

This plots duration on the X-axis and CCI on the Y-axis, separated by output "y".

##**Unsupervised Learning**
###**Principal Components Analysis**

To explore better predictors, we used pcths25_over with pctbachdeg25_over to predict median income. "pcths25_over" deals with the percent of county residents ages 25 and over whose highest level of education is high school.
```{r}

```
The variable similarly produces a percentage of current predictions of around 77% but does not predict the "> 50,000" bucket well.
```{r, message=FALSE}

```

###**K Means Clustering**

Since only considering the education did not seem to produce good predictions, we decided to include percentage of married people because it seemed to be somewhat correlated to median income. So we predicted using the variables pctmarriedhouseholds, pctbachdeg25_over, and pcths25_over.
```{r}


```

This produces a more accurate prediction for both buckets, "<= 50,000" and "> 50,000". The percentage of correct predictions is around 82%.


```{r, message=FALSE, warning=FALSE}

```



