---
title: "Data Analytics Project 2: Credit Card Default"
author: "Team 7"
resource_files:
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
require(data.world)
require(ggplot2)
require(shiny)
require(dplyr)
require(ISLR)
require(MASS)
require(vcd)
library(class)
knitr::opts_chunk$set(echo = TRUE)
```
  
## **R Session Info**  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
sessionInfo()
```

## **Github Link** 
https://github.com/jdramosf/cs329eda

## **Connecting to data.world** 
```{r, message=FALSE, warning=FALSE}
source("data.R", echo = FALSE)
```

## **Introduction** 
This document contains an analysis on the credit card default rate of individuals from Taiwan. We used different statistical analysis methods to predict the probability of an individual to default on their credit card payments. The dataset was obtained from the UCI Machine Learning Repository and contains 30000 records. 

## **Dataset -  Credit Card Default**
```{r, warning=FALSE}
source("Plots.R", echo = FALSE)
```

```{r, warning=FALSE}
source("dap2.R", echo = FALSE)
```

```{r, warning=FALSE}
renderDataTable({
  df[c("limit_bal", "sex", "age", "marriage", "education", "pay_0", "bill_amt1", "pay_amt1", "default")]
})

table(df1$class,df$default)

renderPlot({
  plot(lda.fit)
})
```
